---
title: 'Reproducible Research: Peer Assessment 1'
author: "Jeff Witt"
date: "April 16, 2017"
output: 
  html_document: 
    keep_md: yes
---


## Loading and preprocessing the data

```{r, warning=FALSE}
library(ggplot2) 
```

```{r}
if(!file.exists('activity.csv')){
  unzip('activity.zip')
}
activityData <- read.csv('activity.csv',stringsAsFactors=FALSE)
```


convert date character vector to date

```{r}
activityData$date<- as.Date(activityData$date, format = "%Y-%m-%d")
```


## Total number of steps taken per day
Histogram of the total number of steps taken each day

```{r} 
stepsByDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
qplot(stepsByDay, xlab='Total steps per day', binwidth = 500)
```


##Mean & Median number of steps taken each day
```{r}
stepsByDayMean <- mean(stepsByDay)
stepsByDayMedian <- median(stepsByDay)
```


##Time series plot of the average number of steps taken
```{r}
  averages <- aggregate(x = list(steps = activityData$steps), by = list(interval = activityData$interval), 
                      FUN = mean, na.rm = TRUE)

ggplot(data = averages, aes(x = interval, y = steps)) + geom_line() + xlab("5-minute interval") + 
  ylab("average number of steps taken")
```

##The 5-minute interval that, on average, contains the maximum number of steps
```{r}
  averages[which.max(averages$steps), ]
```


## Code to describe and show a strategy for imputing missing data
```{r}
missing <- is.na(activityData$steps)
table(missing)
```


## Histogram of the total number of steps taken each day after missing values are imputed
```{r}
newactivityData<- activityData
missing_pos <- which(is.na(activityData$steps))
mean_vec <- rep(mean(activityData$steps, na.rm=TRUE), times=length(missing_pos))
newactivityData[missing_pos, "steps"] <- mean_vec

newstepsByDay <- tapply(newactivityData$steps, newactivityData$date, sum, na.rm=TRUE)
qplot(newstepsByDay, xlab='Total steps per day', bins = 50)
```

## Mean and Median with adjustment in missing data
```{r}
mean(newstepsByDay)
median(newstepsByDay)
```

##Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
newactivityData$dateType <-  ifelse(as.POSIXlt(newactivityData$date)$wday %in% c(0,6), 'weekend', 'weekday')

avgnewactivityData <- aggregate(steps ~ interval + dateType, data=newactivityData, mean)
ggplot(avgnewactivityData, aes(interval, steps)) + 
  geom_line() + 
  facet_grid(dateType ~ .) +
  xlab("5-minute interval") + 
  ylab("avarage number of steps")
```


